---
title: "simpleRegress"
author: "Ashia Graves"
date: "Oct 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model WEIGHT ~ NECK  this is an example  

```{r}
simp <- lm(WEIGHT~NECK,data=bears)
plot(WEIGHT~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


## Compare all models that predict WEIGHT other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 



```{r}
headlenmodel <- lm(WEIGHT~HEADLEN,data=bears)
plot(WEIGHT~HEADLEN,data=bears)
abline(headlenmodel)
summary.lm(headlenmodel)
```

 Neck         ARsqrd:0.87     RST:43.91  
 AGE          ARsqrd:0.55     RST:81.47  

The age model grapgh shows its points to be far away from the line while the simp grapgh has much closer points to it.

```{r}
headlenmodel <- lm(WEIGHT~HEADLEN,data=bears)
plot(WEIGHT~HEADLEN,data=bears)
abline(headlenmodel)
summary.lm(headlenmodel)
```
 
 Best to worst  
 Neck         ARsqrd:0.87     RST:43.91  
 Head length  ARsqrd:0.69     RST:77.81  
 AGE          ARsqrd:0.55     RST:81.47  
 
 The headlen graghp model is scattered but it is still better than the agemodel. Compariong the adjusted R squared being higher at 0.69  and the age model adjusted R squared is 0.55.
## Model WEIGHT ~ HEADWDTH
```{r}
headlwtmodel <- lm(WEIGHT~HEADWTH,data=bears)
plot(WEIGHT~HEADWTH,data=bears)
abline(headlenmodel)
summary.lm(headlwtmodel)
```
   Best to worst    
   Neck         ARsqrd:0.87     RST:43.91  
 Head length  ARsqrd:0.69     RST:77.81  
 Head Width   ARsqrd:0.60     RST:76.41  
 AGE          ARsqrd:0.55     RST:81.47  
 
 Headwdthmodel would have a high residual slandered error of 76.41 being lower than age models 81.47 but now low enogh for headlenght


```{r}
lengthmodel <- lm(WEIGHT~LENGTH,data=bears)
plot(WEIGHT~LENGTH,data=bears)
abline(lengthmodel)
summary.lm(lengthmodel)
```
 Best to worst    
 Neck         ARsqrd:0.87     RST:43.91  
Length        ARsqrd:0.74     RST:61.83  
 Head length  ARsqrd:0.69     RST:77.81  
 Head Width   ARsqrd:0.60     RST:76.41  
 AGE          ARsqrd:0.55     RST:81.47  
  
lenghthmodel is 0.74 headlen model is lower being 0.69 but simp is better being 0.87
















